### 2021-06-30 letter/responseLetter.tex

\end{itemize}

{\bf Reviewer 3}

\begin{itemize}

{\color{blue}{\bf Answer.} Reviewer 3 suggested several improvements to the article. To mitigate the improvements pointed, we reviewed the current version of the paper to better clarify Section 1 (Introduction) and Section 3 (Study Settings). We also present more details on FastDroid and Taint Analysis in Section 2 (Background and related work), and we have revised the entire text, correcting several errors that have gone unnoticed.}

@@

\vspace{0.2cm}

\item It is not clear how this paper can be considered an extension of the DroidXP paper presented in SCAM 2020.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} In the Section 1 (Introduction) we rephrase paragraphs 4 to better clarify how our paper is an extension of DroidXP paper from SCAM 2020.}

@@

\vspace{0.2cm}

\item ``We presented and evaluated DroidXP in a conference paper [11]''
What is DroidXP? It should be introduced to a reader who is not familiar.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} We agree. To mitigate this issue, we introduced better DroidXP at Section 1 (Introduction), paragraph 4.}

@@

\vspace{0.2cm}

\item Section 3.1, the help text is not adding much value

\vspace{0.2cm}

{\color{blue}{\bf Answer.} The section 3.1 presented our Benchmark tool DroidXP. The tool description was important for the paper since it was helpful for our first study conduction. DroidXP helped us to easily integrate different test generation tools, and understand the effect of the static analysis component at each tool. It was useful for us and can be very useful for other researchers. We reviewed the first paragraph of section 3.1 to clarify the importance of DroidXP for our study conduction. We included references to DroidXP GitHub repository in the Section 1 (Introduction), since it might be very useful for both researchers and practitioner.}

@@

\vspace{0.2cm}

\item Section 4.1 can be written more concisely.
* It is not clear what is the value of the malware discussion between pages 13 and 15. It may be removed.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} Thanks for raising this issue. Actually our purpose was to present the characteristics of malware that have or have not been identified using the mining sandbox approach. Our goal is to provide a lower-level intuition about the classes of malware, the mining sandbox approach is effective, and which ones it fails. We rephrased the section 4.1 at this point, to clarify this issue about malware discussion.}

### 2021-06-30 letter/responseLetter.tex

{\color{blue}{\bf Answer.} We agree. Nonetheless, DroidFax is not part of the dynamic analysis tools. According to Bao et al.~\cite{jamrozikZ16}, their goal on using DroidFax was to instrument the APKs with the intent of collecting calls to sensitive APIs, while executing the dynamic analysis tools. However, the use of DroidFax in their study actually lead to an increase in the performance of dynamic analysis tools for mining sandboxes. We did not change any internal component of the dynamic analysis tools, which indeed might include static analysis components. We reviewed the title to mitigate this problem.}

\end{itemize}

{\bf Reviewer 3}

 {\color{blue}{\bf Answer.} Reviewer 3 suggested several improvements to the article and clarifications. Thanks for that. We reviewed the current version of the paper to better clarify Section 1 (Introduction) and Section 3 (Study Settings). We also present more details about DroidXP, FastDroid and Taint Analysis in Section 2 (Background and Related Work), and we have revised the entire text, correcting several errors that have gone unnoticed.}

@@

\begin{itemize}

\item It is not clear how this paper can be considered an extension of the DroidXP paper presented in SCAM 2020.

\vspace{0.2cm}

{\color{blue}{\bf Answer.} Thanks for raising this question. In our previous paper,
  the DroidXP's evaluation revealed a possible
overestimation in the performance of the dynamic analysis tools as
reported in the Bao et al. work~\cite{DBLP:conf/wcre/BaoLL18}---which in the end motivated us to
conduct the non-exact replication of that study. In this current paper, 
we extended DroidXP with a new feature to enable / disable the
static analysis component of DroidFax and integrated an additional
tool (Joker), which allowed us
to reproduce the Bao et al. research by means of a set of new experiments
that reveal the actual performance of the dynamic analysis tools.
In addition, we investigate in this paper the effect of using
taint analysis to complement the mining sandbox approach for
malware identification. In the current version of this paper
we better explain the paper's extension.}

@@

\vspace{0.2cm}

\item ``Another approach to rooting'' What is rooting?

\vspace{0.2cm}

{\color{blue}{\bf Answer.} Rooting is a process performed on Android devices that allows
  users of the Android mobile operating system, to execute privileged commands that are typically unavailable.
  The term comes from the fact that Android system is based on Linux, whose superuser (root) has access to administrative permissions on Linux. Since the ``rooting'' term is not well known, we reviewed the text to avoid its use. Thanks for your comment. }

@@

\vspace{0.2cm}

\item What are sensitive behaviors? ``showed that DroidBot is able to trigger an amount of sensitive behaviors''

\vspace{0.2cm}

{\color{blue}{\bf Answer.} We used the term ``sensitive behaviors'' to represent calls to APIs that access private information of an Android user. Access to a sensitive API (such as location or address book) might lead to privacy issues and the leakage of confidential data. We agree that this terminology is not so precise, so we prefer to use the term ``access to sensitive'' APIs through the paper, instead of sensitive behavior.}

@@

\vspace{0.2cm}

\item Section 4.1 can be written more concisely. It is not clear what is the value of the malware
  discussion between pages 13 and 15. It may be removed.

@@

\vspace{0.2cm}

{\color{blue}{\bf Answer.}Thanks for raising this issue. Actually, our goal in this part of the paper is to characterize the malwares that have or have not been identified using the mining sandbox approach. We then provide a lower-level intuition about the classes of malware the mining sandbox approach is effective to identify and which ones it fails. We reviewed Section 4.1 to motivate this discussion better. Nonetheless, we can remove this part of the paper if the reviewers agree that we should remove it.}

### 2021-09-13 background.tex
final revision
last review on the response letter
Remove Krishna comments
change Works to Work
Final review before publication

\section{Background and related works}

%\kn{This section starts abrupt. We should have a line that says what concepts we introduce on a high level. Also, I feel "sandbox" itself should be one of the backgrounds and have a textbf after the introductory line.}

@@

\subsection{Mining Android Sandbox}

@@

The main purpose of a test generator tool is to program crashes or bugs in general. Nonetheless, it is also possible to use test generator tools to explore program behavior (dynamic analysis), and thus assist in the task of building sandboxes. Regarding test generator tools used for mining Android sandboxes, Jamrozik et al~\cite{DBLP:conf/icse/JamrozikZ16} proposed DroidMate, a test generation tool that implements a pseudo-random graphical user interface (GUI) exploration strategy, and was the first approach to leverage test generation to extract sandbox rules from apps. Li et al.~\cite{DBLP:conf/icse/LiYGC17} proposed DroidBot, a test generator tool that explores sensitive resources access from Android apps, following a model-based exploration strategy. In their work, the authors present a comparison between DroidBot and Monkey~\cite{Monkey} regarding malware analysis and show that DroidBot can trigger several sensitive calls more often than Monkey. Sensitive calls in the Android context occurs when an Android app functionality can result in accessing or leaking of Android users' sensitive data. Examples of sensitive calls access user location or network information. Humanoid is another test generator tool for Android~\cite{DBLP:conf/kbse/LiY0C19}---actually a DroidBot evolution. It is also a GUI test generator that learning how humans interact with Android apps. In contrary to random input generators, Humanoid uses a learned model to generate human-like test inputs, and prioritize the possible interactions on a GUI, according to their importance.

>>>>

\section{Background and Related Work}

@@

\subsection{Mining Android Sandbox}\label{sec:android-sandbox}

@@

The main purpose of a test generator tool is to program crashes or bugs in general. Nonetheless, it is also possible to use test generator tools to explore program behavior (dynamic analysis), and thus assist in the task of building sandboxes. Regarding test generator tools used for mining Android sandboxes, Jamrozik et al.~\cite{DBLP:conf/icse/JamrozikZ16} proposed DroidMate, a test generation tool that implements a pseudo-random graphical user interface (GUI) exploration strategy, and was the first approach to leverage test generation to extract sandbox rules from apps. Li et al.~\cite{DBLP:conf/icse/LiYGC17} proposed DroidBot, a test generator tool that explores sensitive resources access from Android apps, following a model-based exploration strategy. In their work, the authors present a comparison between DroidBot and Monkey~\cite{Monkey} regarding malware analysis and show that DroidBot can trigger several sensitive calls more often than Monkey. Sensitive calls in the Android context occurs when an Android app functionality can result in accessing or leaking of Android users' sensitive data. Examples of sensitive calls access user location or network information. Humanoid is another test generator tool for Android~\cite{DBLP:conf/kbse/LiY0C19}---actually a DroidBot evolution. It is also a GUI test generator that learning how humans interact with Android apps. In contrary to random input generators, Humanoid uses a learned model to generate human-like test inputs, and prioritize the possible interactions on a GUI, according to their importance.
